### 2024-09-27 paper/paper.qmd
Completed initial complete draft of the paper.
Added data analysis, figures, descriptions, reference and appendix to the paper. Finalized data cleaning and visualization sections. Ready for review.

---
title: "Analysis of Shootings and Firearm Discharges in Toronto"
author: Jin Zhang
thanks: "Code and data are available at: https://github.com/KrystalJin1/Shootings-Firearm-Discharges-Research.git"
date: September 24, 2024
date-format: long
abstract: "In this paper, we study reported shootings and gun incidents in Toronto from 2018 to 2024. Our analysis identifies key trends over time and in different neighborhoods, showing areas of higher risk, seasons, and times of day. The data indicates a possible relationship between the number of incidents and different areas of Toronto, as well as the time of day. This study provides important insights that will help the Toronto Police Service improve security measures and effectively target high-risk locations with increased police presence."
format: pdf
toc: true
toc-depth: 3
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
```

# Introduction
We used @citeR to analyze the data and plot the graphs for this paper. The folder structure for this paper follows the startup folder created by @rohan. It is very helpful in keeping everything organized. I also referenced many of the examples and techniques provided by @alexander2023telling, which helped me understand how to visualize the data and communicate the findings effectively. We used @citedplyr to process and transform the dataset for this paper efficiently, which made it easier to filter, group, and summarize the data. In addition, the @citetidyverse packages is essential for simplifying data cleaning and analysis, providing a consistent framework for working with data in R. These tools help to organize and present data efficiently.

Shootings and gun discharges have become a growing concern in Toronto. The increase in gun-related incidents not only threatens public safety, but also stresses policymakers, law enforcement and local communities. Understanding the causes of these events, as well as high-risk areas, is essential to developing strategies to fight gun violence. Detailed analysis of data related to shooting incidents can reveal important insights to help identify high-risk areas and times and provide solutions to reduce violence.

Despite the growing concern about gun violence, the public remains poorly informed about how specific factors such as time and place influence gun violence. This requires a more comprehensive analysis of the data. In this paper, I analyze Toronto's reported shooting and gun discharge incidents between 2018 and 2024 and provide the public with relevant information in a brief, easy-to-read chart. We utilize “year” and “month” variables to reveal trends in the number of victims over time and utilize latitude and longitude data to map the density of victims in order to determine which regions and time periods are at a higher risk of having a shooting incident. In addition, we explore the relationship between the hours of the day and the number of related cases.

Data on shooting incidents over the past two decades show fluctuations in the number of victims from year to year, but the number of the incidence has increased in recent years. The frequency of shootings is significantly higher during the summer months, and the incidents tend to occur at night, especially between 9:00 p.m. and 2:00 a.m. The data shows that the most dangerous areas are centered in the downtown area. The data show that the most dangerous areas are concentrated in the city centers. These findings highlight specific times and places that may require increased attention in terms of crime prevention and security measures.

The rest of the paper is structured as follows: in @sec-data, we describe the source of the dataset and its context in detail. Next, we show our findings through various types of graphs with visualizations. After each chart, we discuss each graph's formation, the significance of the findings, and finally summarize key observations. Finally, there is an appendix and references. By the end of this paper, the reader will have a clearer understanding of trends in gun violence in Toronto and how this data can be utilized to improve safety measures throughout the city.

# Data {#sec-data}

## Source and Measurement
The dataset analyzed in this study, the "Shootings & Firearm Discharges" dataset, was retrieved from Open Data Toronto [@citeOpenDataToronto]. This dataset was chosen because it provides detailed information on shooting-related incidents reported to the Toronto Police Service from 2004 to 2024, covering various details like time, location, and number of victims. The information about time and location is particularly useful for identifying trends and pinpointing high-risk periods and locations within the city.This dataset is excellent to help us achieve our research goals.

```{r}
#| eval: true 
#| label: fig-hist-bar
#| fig-cap: "Visualization of Events and People Involved in Shootings and Firearm Discharges"
#| fig-subcap: ["Number of Events by Year", "Number of Events by Time Range", "Total People Involved"]
#| layout-ncol: 3
#| echo: false
#| warning: false
#| message: false

cleaned_data <- read_csv(here::here("data","analysis_data","cleaned_shootings_data.csv"))

# Panel (a) - Histogram for year
ggplot(cleaned_data, aes(x = year)) +
  geom_histogram(binwidth = 1, fill = "steelblue", color = "white") +
  labs(x = "Year", y = "Count", title = "Number of Events by Year") +
  theme_minimal()

# Panel (b) - Bar plot for time_range

# Convert 'time_range' to a factor and order it from morning to night
cleaned_data$time_range <- factor(cleaned_data$time_range, levels = c("Morning", "Afternoon", "Evening", "Night"))

ggplot(cleaned_data, aes(x = time_range, fill = time_range)) +
  geom_bar(stat = "count") +
  labs(x = "Time Range", y = "Count", title = "Number of Events by Time Range") +
  scale_fill_manual(values = c("steelblue", "lightblue", "darkblue", "blue")) +
  theme_minimal()

# Panel (c) - Histogram for total_people
ggplot(cleaned_data, aes(x = total_people)) +
  geom_histogram(binwidth = 1, fill = "steelblue", color = "white") +
  labs(x = "Total People Involved", y = "Count", title = "Total People Involved") +
  theme_minimal()

```
@fig-hist-bar provides an overall view of the number of shooting incidents, victims involved, and time distribution in Toronto from 2004 to 2024.  @fig-hist-bar-1 shows the number of incidents by year**, using the variable “year” to plot the distribution of incidents over time. The results suggest that the number of incidents has increased significantly in recent years, possibly as a result of poor regulation.

@fig-hist-bar-2 represents the **number of events by time range**, based on the `time_range` and 'total_people' variable. This plot reveals that nighttime is a particularly dangerous period for shootings and firearm discharges.

@fig-hist-bar-3 describes the **total** number of people involved in each incident using the “total_people” variable. The analysis showed that most incidents resulted in fewer than three victims, suggesting that although shootings are frequent, the number of people involved in each incident is usually small.

\newpage
```{r}
#| label: fig-deaths-injuries-year
#| fig-cap: Trend of Deaths and Injuries by year
#| echo: false
#| warning: false
#| message: false

yearly_data <- cleaned_data %>%
  group_by(year) %>%
  summarise(total_people = sum(total_people, na.rm = TRUE))

ggplot(yearly_data, aes(x = year, y = total_people)) + 
  geom_line(color = "steelblue", size = 1) + 
  geom_point(color = "steelblue", size = 2) + 
  labs(x = "Year", y = "Total Number of People Involved") + 
  theme_minimal()
```
@fig-deaths-injuries-year shows the trend of the total number of people involved in incidents, including deaths and injuries, over the years from 2004 to 2024. In the early 2000s, the number of people involved was quite high, reaching above 250 in 2004. After that, there is a general decrease, especially around 2010, where the number drops to around 150. However, after 2015, the numbers start to rise again, peaking at over 250 in 2019. After 2020, the number of people involved starts to decrease sharply again. This figure shows how the total number of incidents has changed over time, with clear periods of increase and decrease.

```{r}
#| label: fig-deaths-injuries-month
#| fig-cap: Distribution of Deaths and Injuries by Month
#| echo: false
#| warning: false
#| message: false

# Convert 'month' to a factor and order it chronologically
cleaned_data$month <- factor(cleaned_data$month, levels = month.name)

# Plot Bar chart to show the total number of accidents in different months
ggplot(cleaned_data, aes(x = month, y = total_people, fill = month)) +
  geom_bar(stat = "identity", fill = "steelblue", alpha = 1) +
  labs(x = "Month", y = "Total Number of People Involved") +
  coord_flip() +
  theme_minimal()
```
@fig-deaths-injuries-month shows the total number of shootings and gun discharges by month. The chart uses the “month” variable to show the distribution of incidents throughout the year. From the graph we can observe that the number of shooting victims peaks in the summer months, especially in July and August. This shows that shootings tend to occur more frequently in the warmer months, which may be due to an increase in outdoor activities. Winter months have significantly fewer incidents, suggesting that there may be a seasonal pattern to gun violence in Toronto.

```{r}
#| label: fig-deaths-injuries-time
#| fig-cap: Distribution of Deaths and Injuries by Time Range
#| echo: false
#| warning: false
#| message: false

# Convert 'time_range' to a factor and order it from morning to night
cleaned_data$time_range <- factor(cleaned_data$time_range, levels = c("Morning", "Afternoon", "Evening", "Night"))

# Plot Bar chart to show the total number of accidents in different time ranges
ggplot(cleaned_data, aes(x = time_range, y = total_people, fill = time_range)) + 
  geom_bar(stat = "identity", alpha = 1) + 
  scale_fill_manual(values = c("steelblue", "bisque3", "darkslategray3", "lightpink3"), name = "Time Range") + 
  labs(x = "Time Range", y = "Total Number of People Involved") + 
  theme_minimal()
```

@fig-deaths-injuries-time shows the total number of people involved in shootings and gun incidents at different times of the day. The chart is divided into four time periods: morning, afternoon, evening, and night. The data shows that the majority of incidents occur in the **Evening** and **Night**, where more than 1,500 people are affected. **Afternoon** shows a moderate incidence of incidents, while **morning** has the lowest number of people affected by incidents. This pattern suggests that the risk of shooting increases significantly later in the day, particularly at night, and that heightened attention may be required during these times.

```{r}
#| label: fig-shootings-by-hour
#| fig-cap: "Shootings Distribution by Hour of Day"
#| echo: false
#| warning: false
#| message: false

hourly_data <- cleaned_data %>%
  filter(!is.na(hour)) %>%  # Remove rows where hour contains NA
  group_by(hour) %>%
  summarise(total_events = n()) #Counts the number of rows (events) for each hour group.

# Plot polar bar chart
ggplot(hourly_data, aes(x = as.factor(hour), y = total_events)) +
  geom_bar(stat = "identity", fill = "steelblue") +  # Fill bars with blue color
  coord_polar(start = 0) +  # Convert to polar coordinates
  labs(x = "Hour of Day", y = "Number of Shootings") +
  theme_minimal()

```
@fig-shootings-by-hour shows the number of shooting incidents that occurred during different hours of the day. Most shootings happen between **9 PM (21:00)** and **midnight (23:00)**, where the number of incidents is higher than at other times, reaching above 500 incidents at 23:00. The number of incidents decreases in the morning hours, with fewer shootings happening between **6 AM (6:00)** and **12 PM (12:00)**. This graph shows the hours have higher risk for shootings  during the day.

```{r}
#| label: fig-shootings-by-map
#| fig-cap: "Map of Shootings Highlighted by Risk Level"
#| echo: false
#| warning: false
#| message: false

# Filter out unreasonable latitude and longitude values (for Toronto region)
map_data <- cleaned_data %>%
  filter(Longitude < -70 & Longitude > -90, Latitude > 40 & Latitude < 45)

# Create a Risk_Level column, assuming 'DEATH' and 'INJURIES' columns indicate casualties
map_data <- map_data %>%
  mutate(Risk_Level = case_when(
    death > 0 ~ "High Risk",  # Define as high risk if there is a death
    injuries > 2 ~ "High Risk",  # Define as high risk if there are more than 2 injuries
    TRUE ~ "Low Risk"  # Otherwise, define as low risk
  ))

# Plot the map using the newly created Risk_Level column
ggplot(map_data, aes(x = Longitude, y = Latitude, color = Risk_Level)) +
  geom_point(alpha = 0.8, size = 0.7) + 
  scale_color_manual(values = c("High Risk" = "brown2", "Low Risk" = "steelblue")) +  # Manually set colors
  labs(title = "Map of Incidents Highlighted by Risk Level",
       x = "Longitude", y = "Latitude",
       color = "Theft Risk Level") +  
  theme_minimal() +
  theme(panel.background = element_rect(fill = "white"),  
        legend.position = "right",  # Place legend on the right
        plot.title = element_text(hjust = 0.5, face = "bold"))

```
@fig-shootings-by-map shows the geographic distribution of shootings in Toronto, categorized by risk level. Events with more than two deaths or injuries are defined as high risk. The map uses blue dots to indicate low-risk incidents and red dots to indicate high-risk incidents. The map shows that most incidents are categorized as low risk. High-risk incidents tend to occur in the central areas of Toronto, where there is a higher density of dots. This map helps us to understand where the shooting incidents are high and to be safe when traveling.

\newpage

\appendix

# Appendix {-}

## Appendix A: Data Cleaning Process

### Initial Data Cleaning
The first step involves loading the original dataset, normalizing and simplifying the column names. This makes the data easier to analyze. Only the most important variables were kept, focusing on information such as the time, location and outcome (e.g., deaths and injuries) of the event.

### Renaming Variables
Some variables were renamed to make the dataset easier to interpret. For example, `occ_date` was renamed to `date` and `occ_year` was renamed to `year`. Renaming is an important step in ensuring that we label variables clearly.

### Creating the total_people Variable
A new variable “total_people” was created by adding the values from the `deaths` and `injuries` columns. If one of the columns had missing data (i.e., an NA value), it was replaced with a zero. This adjustment provides a more complete and accurate view of the total number of people affected in each incident.

### Splitting the Geometry Column
The `geometry` column containing the geographic coordinates is split into two columns: `longitude` and `latitude`. This conversion makes it easier to analyze and plot the

### Saving the Cleaned Dataset
After cleaning and transforming the data, the final version was saved in a new file.

\newpage
# References